---
title: 'Homework 1: Data manipulation'
author: "IE 0015"
date: "February 14, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

For the homework assignments, code and narrative should be written using R Markdown.  Process maps can be written on separate pieces of paper. You should render the R Markdown as PDF or Word, print the output, and staple with the process maps for the complete submission.

In the `tidyr` package there are two Tubeculosis data sets `who` and `population` (you used them in Lab 3).  Look at the `tidyr` help files (R Studio -> Packages tab -> tidyr) to see what each contains. To load and view the last few rows of the datasets, you should enter:

```{r, warning=FALSE}
library(tidyr)
library(dplyr)
library(magrittr)
library(stringr)
data(who)
tail(who)
```
```{r}
data(population)
tail(population)
```

1. [4] Look the `who` and `population` data sets from the `tidyr` package (the information files are in the `tidyr` package help files).  Make the `who` dataset tidy data (Note: see lab 3) with the following columns:  Country, iso2, year, subgroup, cases.  For `subgroup`, the value should be what are currently column headings. (in lab 3 you used `agerange`).

```{r}
theWho <- who %>% gather("new_sp_m014",  "new_sp_m1524","new_sp_m2534", "new_sp_m3544", "new_sp_m4554", "new_sp_m5564", "new_sp_m65", "new_sp_f014", "new_sp_f1524", "new_sp_f2534", "new_sp_f3544", "new_sp_f4554", "new_sp_f5564", "new_sp_f65",  
"new_sn_m014",  "new_sn_m1524", "new_sn_m2534", "new_sn_m3544", "new_sn_m4554", "new_sn_m5564","new_sn_m65",   "new_sn_f014",  "new_sn_f1524", "new_sn_f2534", "new_sn_f3544", "new_sn_f4554","new_sn_f5564", "new_sn_f65",   
"new_ep_m014",  "new_ep_m1524", "new_ep_m2534", "new_ep_m3544","new_ep_m4554", "new_ep_m5564", "new_ep_m65",   "new_ep_f014",  "new_ep_f1524", "new_ep_f2534","new_ep_f3544", "new_ep_f4554", "new_ep_f5564", "new_ep_f65",   
"newrel_m014",  "newrel_m1524","newrel_m2534", "newrel_m3544", "newrel_m4554", "newrel_m5564", "newrel_m65",   "newrel_f014","newrel_f1524", "newrel_f2534", "newrel_f3544", "newrel_f4554", "newrel_f5564", "newrel_f65", key = "subgroup", value = "cases")

tidyWho <- theWho
tidyWho$iso3<-NULL

tail(tidyWho)
```

2.  [10] In the subgroup column, there are actually three pieces information: the mode of diagnosis,  gender, and the age range.
  a. Draw a process map for the steps to take to extract age range, diagnosis mode, and gender by year from the original column names. Note: in a process map, each function in (b) should correspond to a block on the process map.
  b. In R, create three new columns 'agerange', 'gender', 'diagnosismode' and populate them based on the current value of subgroup. (note: 'agerange' is going to be the hardest of the three, I suggest starting with 'gender'). Print out the fifst 10 rows of Country, iso3, year, agerange, gender, diagnosismode, cases.  Note: because of the size of the dataset, start by making a dataframe with the rows of `tidywho` from question 1 where the year is after 2010 (i.e. 2010-2014).
  
```{r}
subgroup_No_New <- data_frame(subgroup = c(str_sub(tidyWho$subgroup, 4)))

theWho$gender <- c(str_extract(theWho$subgroup,"f|m"))
theWho$diagnosismode <- c(str_extract(subgroup_No_New$subgroup,"[a-zA-Z0-9]{2,3}"))
theWho$agerange <- c(str_extract(subgroup_No_New$subgroup,"\\d{1,5}"))

theNewWho = data_frame(Country = c(theWho$country), iso3 = c(theWho$iso3), Year = c(theWho$year), AgeRange = c(theWho$agerange), Gender = c(theWho$gender), DiagnosisMode = c(theWho$diagnosismode), Cases = c(theWho$cases) )

head(theNewWho,10)
```


```{r}
theWho2011 <- theWho %>% filter(year>2010)
```

3.  [6] You need a table of rates of annual diagnosis by country for the last 3 years (2011-2013).
  a. Draw a process map that creates a table with the total rate (per 100,000) of new diagnosis for each country for each year by diagnosis mode. Note: it this table should be tidy.
  b. Implement the procedure, then you should have a table with: Country, iso3, diagnosismode, rate.  Note that this requires combining the data elements for gender and age ranges.
  c. Display the last 15 rows (5 countries).
  
```{r}
theWho2011$rates <- c(theWho2011$cases/100000)
groupedWho<-group_by(theWho2011,country, year, diagnosismode, iso3)
summarizedWho2011 <- summarize(groupedWho, sum(cases))

ratesWho2011 <- data_frame(Country = c(summarizedWho2011$country), iso3 = c(summarizedWho2011$iso3), DiagnosisMode = c(summarizedWho2011$diagnosismode), Rate = c(summarizedWho2011$`sum(cases)`)/100000)

tail(ratesWho2011, 15)
```
  
  
  