---
title: 'Lab 10: Regression'
author: "IE 0015 Information Systems Engineering"
date: "March 29/30, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

This lab will use the Titanic data. You should modify the `read.csv` function call to correspond to your system.

```{r loadlibraries,  warning=FALSE}
options(digits = 3, scipen=4)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(magrittr)
library(modelr, quietly = TRUE, warn.conflicts = FALSE)

titanic <- read.csv("~/Downloads/titanic3.csv",
                    header=TRUE, sep=",")
# From Lab on working with characters and regular expressions
getTitle <- function(data) {
  title.dot.start <- regexpr("\\,[A-Z ]*\\.", data$name, TRUE)
  title.comma.end <- title.dot.start+ attr(title.dot.start, "match.length")
  data$Title <- substr(data$name, title.dot.start+2, title.comma.end-1)
  return (data$Title)
}  
titanic$Title <- getTitle(titanic)
unique(titanic$Title)  
```


# 1. Multiple linear regression (4 pts)

You think that the age is related to the following:

  i.  Sex
  ii.  Pclass

a.  Remove lines in the data that do not have an age
b.  Develop a multiple linear regression model for the age based on these characteristics.
c.  Display the summary and the analysis of variance tables for this model. 
d.  Display diagnostics plots. Discuss the assumptions of regression models.
e.  Do you think this is an adequate model?

```{r}
titanic <- filter(titanic, age > 0) #removes lines with no age
regresTitanic <- lm(age ~ sex + pclass, data=titanic)
summary(regresTitanic)
anova(regresTitanic)
plot(regresTitanic)
```
D.) NORMAL Q-Q IS SUPOSED TO BE LINEAR IN THE MIDDLE
    Y SHOULD BE A COMBINATION OF MU AND RANDOM ERROR
    RANDOM ERROR SHOULD BE RANDOMLY DISTRIBUTED
    
E.) THIS IS AN ACCURATE MODEL BECAUSE THE ERRORS ARE RANDOMLY DISTRIBUTED AND      THE PLOT Q-Q SHOWS THAT OUR PREDICTIONS ARE WELL ALIGNED WITH THE ACTUAL       DATA.
  
# 2. Adding to the model. (4 pts)

There are two variables that are generally present but are not currently in the model: fare and Title. 

a. Plot the residuals of your model against each of these and choose which value should be added to the model.
b. Create the linear regression model that adds the new variable and print the summary, anova table, and diagnostic plots.
c. Compare the new model to the original.

```{R}
ggplot(titanic, aes(x = regresTitanic$residuals, y=titanic$fare)) + geom_point()

ggplot(titanic, aes(x = regresTitanic$residuals, y=titanic$Title)) + geom_point()

regres2Titanic <- lm(age ~ sex + pclass + Title, data=titanic)
summary(regres2Titanic)
anova(regres2Titanic)
plot(regres2Titanic)
```
THE NEW MODEL LOOKS SLIGHTLY WORSE THAN THE OTHERS AS THE Q-Q GRAPH DEVIATES SIGNIFICANTLY FROM THE LINE WITH SLOPE 1. ALSO, THE RESIDUAL GRAPHS HAVE LINES WHERE THE SLOPE IS NOT A CONSTANT 0.The p-value for the Title is pretty small so the Title category is significant.

3.  Use the final model from problem 2 to predict the ages of the individuals listed below. (2 points)  

Hint: look up `predict` in help.

Note: remember to explain the output.


```{r predictpassenger}
predicttitanic <- titanic[c(13, 51), c("pclass", "name", "sex", "sibsp", "embarked", "Title")]
predicttitanic

predict(regres2Titanic, newdata = predicttitanic)
```
USING THE DATA GIVEN, THE MODEL ESTIMATES THE AGE OF A PERSON WITH THE GIVEN PCLASS, SEX, AND TITLE.


